---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r}
        rawData <- read.csv(unzip("activity.zip"))
```
## What is mean total number of steps taken per day
```{r}
        library(dplyr)
        library(ggplot2)
        sum_by_day <- rawData %>%
          group_by(date) %>%
            summarise(steps_taken_by_day = sum(steps, na.rm=TRUE)) %>%
          ungroup()
        mean_steps <- mean(sum_by_day$steps_taken_by_day, na.rm=TRUE)
        median_steps <- median(sum_by_day$steps_taken_by_day, na.rm=TRUE)
        
        sum_by_day
        ggplot(sum_by_day, aes(x=steps_taken_by_day))+ geom_histogram(binwidth = 1000) + xlab("Total Steps by day") + ylab("Frequency") + scale_y_continuous(breaks = scales::breaks_width(2)) + ggtitle("Steps per day Histogram")
        mean_steps
        median_steps
```
The mean steps is 9354.23 and the median steps is 10395 of the total number of steps taken per day.

## What is the average daily activity pattern?
```{r}
        mean_steps_by_interval <- rawData %>%
          group_by(interval) %>%
            summarise(mean_steps = mean(steps, na.rm=TRUE)) %>%
          ungroup()
        ggplot(mean_steps_by_interval, aes(x=interval, y=mean_steps)) + geom_line() + xlab("Interval") +     ylab("Average number of steps taken") + ggtitle("Daily active pattern")
        max_val <- max(mean_steps_by_interval[[2]], na.rm = TRUE)
        max_interval <- mean_steps_by_interval[mean_steps_by_interval[[2]] == max_val,1]
        max_interval
```
5-minute interval contains the maximum number is 835.

## Imputing missing values
```{r}
        sum(!complete.cases(rawData))
```
The total number of missing values is 2304.
I replaced NA steps with the mean steps calculated for each intervals.
```{r}
        rawData_imputed <- rawData %>%
          group_by(interval) %>%
          mutate(
            steps_imputed = if_else(
              is.na(steps), mean(steps, na.rm=TRUE),
              steps
            )
          ) %>%
          ungroup()
        
        sum_by_day_imputed <- summarise(group_by(rawData_imputed,date),steps_taken_by_day_imputed = sum(steps_imputed, na.rm=TRUE))
        mean_steps_imputed <- mean(sum_by_day_imputed$steps_taken_by_day_imputed, na.rm=TRUE)
        median_steps_imputed <- median(sum_by_day_imputed$steps_taken_by_day_imputed, na.rm=TRUE)
        
        ggplot(sum_by_day_imputed, aes(x=steps_taken_by_day_imputed))+ geom_histogram(binwidth = 1000) + xlab("Total Steps by day") + ylab("Frequency") + scale_y_continuous(breaks = scales::breaks_width(2)) + ggtitle("Imputed Dataset Histogram")
        mean_steps_imputed
        median_steps_imputed
```
After imputation, the mean steps is 1.076619 and the median steps is 1.0769.
The imputed mean and median steps are same.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
        rawData$date <- as.Date(rawData$date)
        Data_weekday <- rawData %>%
          mutate(
            is_weekday = !weekdays(date) %in% c("Saturday", "Sunday", "土曜日", "日曜日")
          )
        weekday_true <- subset(Data_weekday, is_weekday == TRUE)
        weekday_false <- subset(Data_weekday, is_weekday == FALSE)
       
        mean_steps_by_interval_weekday <- weekday_true %>%
          group_by(interval) %>%
            summarise(mean_steps = mean(steps, na.rm=TRUE)) %>%
          ungroup()
        ggplot(mean_steps_by_interval_weekday, aes(x=interval, y=mean_steps)) + geom_line() + xlab("Interval") +     ylab("Average number of steps taken") + ggtitle("Weekday")
        
        mean_steps_by_interval_weekend <- weekday_false %>%
          group_by(interval) %>%
            summarise(mean_steps = mean(steps, na.rm=TRUE)) %>%
          ungroup()
        ggplot(mean_steps_by_interval_weekend, aes(x=interval, y=mean_steps)) + geom_line() + xlab("Interval") +     ylab("Average number of steps taken") + ggtitle("Weekend")
        
```
